---
title: "It brings me ggjoy"
author: "@awhstin"
date: 2017-06-15T21:13:14-05:00
categories: ["R"]
titleimg: "joy-1"
---
```{r, message=FALSE, warning=FALSE,include=FALSE}
library(extrafont)
weather.raw<-read.csv('~/Desktop/austin/nebraska-2016.csv',stringsAsFactors = F)
```
Awhile ago I posted about plotting the temperatures of Lincoln Nebraska that was inspired by a FiveThirtyEight article visualization. Well the internet have been abuzz with a new package found on <a href="https://github.com/clauswilke/ggjoy" target="_blank">github</a> by Claus Wilke called <code class="language-r">ggjoy</code>. So I decided to do a quick little post playing with it.

<h3>code</h3>

Once you have the package installed from github, go ahead and download this <a href="https://drive.google.com/file/d/0ByOfjCmqEilLYndpOWJyZXhPVUk/view" target="_blank">csv</a> of the 2016 temperatures in Lincoln, NE. Then we can clean and format it to plot.

```{r, message=FALSE, warning=FALSE}
library(ggjoy)
library(hrbrthemes)
weather.raw$month<-months(as.Date(weather.raw$CST))
weather.raw$months<-factor(weather.raw$month,levels=unique(weather.raw$month))

#scales
mins<-min(weather.raw$Min.TemperatureF)
maxs<-max(weather.raw$Max.TemperatureF)
```

To plot I decided to use the always sharp looking <code class="language-r">hrbrthemes::theme_ipsum</code> and with a few lines we get a really nice looking joy plot. Which for the record is an awesome name, read <a href="https://blogs.scientificamerican.com/sa-visual/pop-culture-pulsar-origin-story-of-joy-division-s-unknown-pleasures-album-cover-video/" target="_blank">this</a>. 

```{r joy, fig.align='center', fig.height=8, fig.path='../', fig.width=10, message=FALSE, warning=FALSE, dev='svg', include=FALSE}
ggplot(weather.raw,aes(x = Mean.TemperatureF,y=months,height=..density..))+
     geom_joy(scale=3) +
      scale_x_continuous(limits = c(mins,maxs))+
  theme_ipsum(grid=F)+
  theme(axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 180, hjust = 1))+
  labs(title='Temperatures in Lincoln NE',
       subtitle='Median temperatures (Fahrenheit) by month for 2016\nData: Original CSV from the Weather Underground')
```
```{r fig.align='center', fig.height=8,fig.width=10}
ggplot(weather.raw,aes(x = Mean.TemperatureF,y=`months`,height=..density..))+
     geom_joy(scale=3) +
      scale_x_continuous(limits = c(mins,maxs))+
  theme_ipsum(grid=F)+
  theme(axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.text.y = element_text(angle = 180, hjust = 1))+
  labs(title='Temperatures in Lincoln NE',
       subtitle='Median temperatures (Fahrenheit) by month for 2016\nData: Original CSV from the Weather Underground')
```